Fast Algorithms for Intersection of Non-matching Grids Using \plucker Coordinates.

### 2016-10-01 paper_intersections.tex

\section{Advancing Front Method}
\label{sec:front_advancing}
add references...
% BIH:
% http://ainc.de/Research/BIH.pdf
% http://diglib.eg.org/handle/10.2312/EGWR.EGSR06.139-149

Consider now a complex mesh of combined dimensions consisting of \emph{components}, which are sets of connected
elements of the same lower dimension (1d, or 2d), in the space of connected 3d elements, which we shall call a \emph{bulk}. 
Obtaining all of component-bulk intersections is done in two phases: firstly, we look for the first two elements intersecting each other (initialization); secondly, we prolong the intersection by investigating neighbouring elements (intersection tracking).

To construct the Advancing front algorithm, we shall need:
\begin{itemize}
    \item \textbf{element connectivity} -- we assume this data is available from mesh preprocessing,
    \item \textbf{Axes Aligned Bounding Boxes (AABB)} -- we construct these in order to decide fastly
                    whether to compute the actual intersection of two elements,
    \item \textbf{Bounding Interval Hierarchy (BIH)} -- we alternatively create BIH above AABB to fastly search 
                    created bounding boxes for two colliding with each other and thus obtaining a candidate pair.
\end{itemize}

The intersection tracking itself can be also seen as a \emph{breadth-first search}
\footnote{Wiki, [online 2016-03-01], \url{https://en.wikipedia.org/wiki/Breadth-first_search}}
algorithm over the BIH, following the component elements.

@@

\paragraph{Initialization}
We start with selecting an arbitrary 1d or 2d element. Then we search the bulk elements, checking for 
a collision of bounding boxes, to create a candidate pair. Using only AABB, we need to iterate over
bulk elements in $O(n)$, $n$ being the number of all elements in our case. Using BIH, we can speed 
up the search to $O(\log n)$ on average. In later case, we are paying the costs in the construction of BIH,
which is a quicksort like algorithm running at $O(n\log n)$ on average.

Now that we have provided the first candidate pair, we can look at the scheme in \fig{fig:prolongation} and see us moving from the green box in the left upper corner. If an intersection exists, we have just started a new component and we can procced to tracking the intersection. Otherwise, we select another 1d or 2d element and start over.

@@

Let us now discuss the advancing front algorithm displayed by the scheme in \fig{fig:prolongation}. The main idea is to compute intersections for a component element with all possible bulk elements, and then move to a~next neighboring component element. For this reason, we define two
queues of candidate pairs: a~\emph{bulk queue} and a~\emph{component queue} (yellow boxes).

On input we consider a candidate pair, for which a non-empty intersection is computed.
Now we look for new candidate pairs among the neighboring elements (the block Prolongation decide).
Therefore, we iterate over the intersection points and further exploit the topological information.
There are 3 possible cases (applies both for 1d-3d and 2d-3d), how the intersection might be prolonged:
\begin{itemize}
    \item \textbf{IP lies on the component element side and inside the bulk element} \\
            We find all the sides of component element in which the IP lies (IP can be at node and connect more sides).
            Next, we find the component neighboring elements over the sides and push all new candidate pairs
            [component neighbor -- current bulk element] into the component queue.
            Note, that there can be more than one neighbor on a side, if the component has branches.
    \item \textbf{IP lies on the component element side and on the surface of the bulk element} \\
            We find all the faces of bulk element in which the IP lies (1 face, or 2 faces (IP on an edge), 
            or 3 faces (IP at a node)). We find the corresponding neighboring bulk elements over the faces and
            push the new candidate pairs [current component element -- bulk neighbor] into the bulk queue and [component neighbor -- bulk neighbor] into the component queue.
    \item \textbf{IP lies inside component element (therefore must be on the surface of bulk element)} \\
            We proceed as in previous case, but we push only [current component element -- bulk neighbor] candidate pairs, since there is no component neighbor.
\end{itemize}

>>>>

\section{Global mesh intersection algorithm}
\label{sec:front_advancing}
Having the algorithms for element-element intersections we can proceed to the mesh intersection algorithm. 
We first compute all component-bulk mesh intersections that is (1d-3d and 2d-3d) and then the 1d-2d and 2d-2d intersections use the 3d mesh
to get the intersection pair candidates.  Element intersections for the particular component and bulk mesh 
are obtained in two phases: firstly, we look for the first two elements intersecting each other (initialization); 
secondly, we prolong the intersection by investigating neighbouring elements (front tracking).
\noteJB{Not sure that "prolong" is the right verb.}

@@

\paragraph{Initialization}
We choose an arbitrary element of the component mesh and check it intersection with all elements of the bulk mesh.
The process is accelerated by computing the axis aligned bounding box (AABB) for every element and use intersection of the bounding boxes
as a fast indicator for possible intersection of the elements. This step takes $O(N)$ time with respect to the number of elements of the bulk mesh $N$.
For the small number of components $k$ and if they are contained inside the bulk mesh the total time of initialization may be still linear $O(kn)$. However,
for more complex cases we organize the bounding boxes of the bulk mesh into the bounding interval hierarchy (BIH) \cite{EGWR:EGSR06:139-149} 
a data structure in principle equivalent 
to R-trees \cite{guttman_r-trees:_1984}, \cite{nam_comparative_2004}. The creation time is $O(N\log(N/n))$ and the lookup time $O(\log(N/n))$ where $n$ is number
of the bulk elements in the leaf nodes of the tree.

\paragraph{Advancing Front Method}
The advancing front algorithm requires and employ the neigbouring information for the elements within the component as well as within the bulk mesh.
It can be viewed as the breadth first search algorithm for a graph which vertices are intersection polygons and edges are sides shared by
two polygons. Side every side of IP is on the boundary of either component or the bulk element we can distinguish bulk and component edges and correspondingly 
the \emph{bulk queue} and the \emph{component queue}.

@@

The scheme of the advancing front algorithm is displayed by the scheme in \fig{fig:prolongation}. 
The main idea is to compute intersections for a component element with all possible bulk elements, and then move to a~next neighboring component element. 
So the bulk queue is emptied before the component queue (see yellow boxes).

On the input we consider a candidate pair provided be the initialization (top-left corner).
Now we look for new candidate pairs among the neighboring elements (the block Prolongation decide).
Therefore, we iterate over the intersection corners and further exploit their topological position on the component element $C$ and the bulk element $B$.
There are 3 possible cases (applies both for 1d-3d and 2d-3d), how the intersection might be prolonged:
\begin{itemize}
    \item \textbf{IC on the boundary of $C$  and inside $B$} \\
            We find all the sides coincident with the \nface of IC on $C$ component element in which the IP lies (IP can be at node and connect more sides).
            Next, we find the component neighboring elements over the sides and push all new candidate pairs
            [component neighbor -- current bulk element] into the component queue.
            Note, that there can be more than one neighbor on a side, if the component has branches.
    \item \textbf{IC on the boundary of both $C$ and $B$} \\
            We find all the faces of bulk element in which the IP lies (1 face, or 2 faces (IP on an edge), 
            or 3 faces (IP at a node)). We find the corresponding neighboring bulk elements over the faces and
            push the new candidate pairs [current component element -- bulk neighbor] into the bulk queue and [component neighbor -- bulk neighbor] into the component queue.
    \item \textbf{IC inside the component element and on the boundary of the bulk element)} \\
            We proceed as in previous case, but we push only [current component element -- bulk neighbor] candidate pairs, since there is no component neighbor.
\end{itemize}
